---
title: "Gene expression heatmaps - practice"
output: html_document
---
Rmd to create heatmaps of differentially expressed gene lists. Two DEG lists (with gene counts):   
1. Temperature DEG and gene count list:
https://raw.githubusercontent.com/RobertsLab/paper-crab/master/analyses/temp_DEGlist-counts.tab 
2. Infection DEG and gene count list: https://raw.githubusercontent.com/RobertsLab/project-crab/master/analyses/2019-infection-degs_genecounts.csv 

ALSO --> after those two heatmaps, I'm making a TON of heatmaps from the 2020 Genewiz individual RNA seq data. 


## Install packages:
(uncomment "install.packages("pheatmap")" if you need to install it)
```{r}
#install.packages("pheatmap")
library(pheatmap)
```

```{r}
library(dplyr)
```

```{r}
#install.packages("RColorBrewer")
library(RColorBrewer)
```

# Temperature comparison (elevated vs decreased) heatmap

Create color scheme and save as object:
(11 --> number of colors in color palette)
```{r}
heatmapRedYelBlu <- RColorBrewer::brewer.pal(11, "RdYlBu")
```

### Read in temperature DEG list with gene counts table (423 rows)
```{r}
temp_counts <- read.delim("../analyses/temp_DEGlist-counts-log.txt", sep = '\t')
head(temp_counts)
```

Subset just the Trinity_ID column and the gene count sample columns (log +1 transformed):
```{r}
matrix_temp <- select(temp_counts, Trinity_ID, X380822_cold_log, X380823_cold_log, X380824_warm_log, X380825_warm_log)
head(matrix_temp)
```

Set Trinity_ID as row names:
```{r}
rownames(matrix_temp) <- matrix_temp$Trinity_ID #set Trinity_ID as rownames
matrix_temp.heatmap <- matrix_temp[,-1] #remove redundant column
head(matrix_temp.heatmap)
```

Subset genes of interest (for practice, just picking random ones)
```{r}
#include_list <- c("TRINITY_DN169_c0_g1_i11", "TRINITY_DN983_c0_g1_i3", "TRINITY_DN25158_c34_g1_i1", "TRINITY_DN5806_c0_g1_i1")
#matrix_temp.heatmap.sub <- matrix_temp.heatmap[include_list, ]
#matrix_temp.heatmap.sub
```

Create heatmap for temperature subset DEG list:
```{r}
pheatmap(matrix_temp.heatmap, cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = FALSE, treeheight_col = 80, legend = TRUE, color = heatmapRedYelBlu, fontsize_col = 20, fontsize_row = 12)
```


# Infection comparison heatmap 

Heatmap color palette: 
```{r}
heatmapPiYG <- RColorBrewer::brewer.pal(11, "PiYG")
```

### Read in infection status DEG list with gene counts table (772 rows)
```{r}
infect_counts <- read.delim("../../project-crab/analyses/2019-infection-degs_genecounts.csv", sep = ',')
head(infect_counts)
```

Subset just the Trinity_ID column and the gene count sample columns:
```{r}
matrix_infect <- select(infect_counts, Trinity_ID, infected_12, uninfected_12, infected_26, uninfected_26)
head(matrix_infect)
```

Set Trinity_ID as row names:
```{r}
rownames(matrix_infect) <- matrix_infect$Trinity_ID #set Trinity_ID as rownames
matrix_infect.heatmap <- matrix_infect[,-1] #remove redundant column
head(matrix_infect.heatmap)
```

Subset genes of interest (for practice, just picking random ones)
```{r}
#include_list <- c("TRINITY_DN1003_c0_g1_i4", "TRINITY_DN7071_c0_g1_i1", "TRINITY_DN1796_c0_g1_i1", "TRINITY_DN1796_c0_g1_i1")
#matrix_infect.heatmap.sub <- matrix_infect.heatmap[include_list, ]
#matrix_infect.heatmap.sub
```

Create heatmap for temperature subset DEG list:
```{r}
pheatmap(matrix_infect.heatmap, cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapPiYG, fontsize_col = 20, fontsize_row = 12)
```

------------------------------------------------------------------------
# HEAT MAPS FOR INDIVIDUAL CRAB RNASEQ DATA
Just making a ton to see if anything is interesting 



