---
title: "DESeq_libraries8-11"
output: html_document
---
Rmd to go through `DESeq2` pipeline. Will check data. Then will perform some DEG analyses. 

Libraries 8-11 info: 
(These were used in the 02-DeSeq-Temperature.Rmd comparing elevated and decreased)     
They are all pooled samples:   
8 --> Day 2, infected, decreased temperature
9 --> Day 2, uninfected, decreased temperature
10 --> Day 2, infected, elevated temperature
11 --> Day 2, uninfected, elevated temperature

```{r}
library(DESeq2)
library(tidyverse)
```


Read in un-normalized counts:
```{r}
pooldf <- read.table("../data/kallisto-0517.isoform.counts.txt", header = T, sep = "\t")
rownames(pooldf) <- pooldf$FeatureID
pooldf <- pooldf[,-1]
head(pooldf)
```

Rename columns so I know which samples are infected and which are uninfected:
```{r}
colnames(pooldf) <- c("uninfected_cold", "infected_cold", "uninfected_warm", "infected_warm")
head(pooldf)
```


## Get DEGs based on infection with temperature as a factor: 
From Multi-factor designs section of: 
http://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#multi-factor-designs 

```{r}
deseq2.colData <- data.frame(condition=factor(c("uninfected", "infected", "uninfected", "infected")), 
                             type=factor(rep("paired-end", 4)),
                             temperature=factor(c("decreased", "decreased", "elevated", "elevated")))
rownames(deseq2.colData) <- colnames(data)
deseq2.dds <- DESeqDataSetFromMatrix(countData = pooldf,
                                     colData = deseq2.colData, 
                                     design = ~ condition + temperature)
```

Check levels of temperature and condition (infection status)
```{r}
levels(deseq2.dds$temperature)
```

```{r}
levels(deseq2.dds$condition)
```

`DESeq2` automatically puts the levels in alphabetical order and the first listed level is the reference level for the factor. 

We want uninfected to be the reference. 
```{r}
deseq2.dds$condition = relevel(deseq2.dds$condition, "uninfected")
levels(deseq2.dds$condition)
```

Cool. 
```{r}
design(deseq2.dds) <- formula(~ temperature + condition)
deseq2.dds <- DESeq(deseq2.dds)
```

Access results:
```{r}
deseq2.res <- results(deseq2.dds)
head(deseq2.res)
```

```{r}
summary(deseq2.res)
```


```{r}
# Count number of hits with adjusted p-value less then 0.05
dim(deseq2.res[!is.na(deseq2.res$padj) & deseq2.res$padj <= 0.05, ])
```

```{r}
inf <- deseq2.res
# The main plot
plot(inf$baseMean, inf$log2FoldChange, pch=20, cex=0.45, ylim=c(-15, 15), log="x", col="darkgray",
     #main="Infection Status  (pval </= 0.05)",
     xlab="mean of normalized counts",
     ylab="Log2 Fold Change")
# Getting the significant points and plotting them again so they're a different color
inf.sig <- deseq2.res[!is.na(deseq2.res$padj) & deseq2.res$padj <= 0.05, ]
points(inf.sig$baseMean, inf.sig$log2FoldChange, pch=20, cex=0.45, col="red")
# 2 FC lines
abline(h=c(-1,1), col="blue")
```

## Now I need to figure out what these results mean: how exactly does temperature impact DEGs based on infection?

```{r}
deseq2.resTemp <- results(deseq2.dds,
                          contrast = c("temperature", "elevated", "decreased"))
head(deseq2.resTemp)
```

```{r}
tmp <- deseq2.resTemp
# The main plot
plot(tmp$baseMean, tmp$log2FoldChange, pch=20, cex=0.45, ylim=c(-15, 15), log="x", col="darkgray",
     #main="Infection Status  (pval </= 0.05)",
     xlab="mean of normalized counts",
     ylab="Log2 Fold Change")
# Getting the significant points and plotting them again so they're a different color
tmp.sig <- deseq2.resTemp[!is.na(deseq2.resTemp$padj) & deseq2.resTemp$padj <= 0.05, ]
points(tmp.sig$baseMean, tmp.sig$log2FoldChange, pch=20, cex=0.45, col="red")
# 2 FC lines
abline(h=c(-1,1), col="blue")
```

```{r}
# Count number of hits with adjusted p-value less then 0.05
dim(deseq2.resTemp[!is.na(deseq2.resTemp$padj) & deseq2.resTemp$padj <= 0.05, ])
```


















