---
title: "DESeq infection"
output: html_document
---
Rmd to get differnetially expressed genes for comparing infected and uninfected crabs (combine temp and time)

Re-doing the script that SR originally did in the new repo (SR's original script: https://github.com/RobertsLab/project-crab/blob/master/scripts/11-Deseq.Rmd)

```{r}
library(DESeq2)
```


```{r}
d19 <- read.table("../data/kallisto-0315.isoform.counts.txt", header = T, sep = "\t")
rownames(d19) <- d19$FeatureID
d19 <- d19[,-1]
head(d19)
```

Rename columns:
```{r}
colnames(d19) <- c("Infected_2", "Infected_17", "Uninfected_2", "Uninfected_17")
head(d19)
```

```{r}
deseq2.colData <- data.frame(condition=factor(c(rep("Infected", 2), rep("Uninfected", 2))), 
                             type=factor(rep("paired-end", 4)))
rownames(deseq2.colData) <- colnames(data)
deseq2.dds <- DESeqDataSetFromMatrix(countData = d19,
                                     colData = deseq2.colData, 
                                     design = ~ condition)
```


```{r}
deseq2.dds <- DESeq(deseq2.dds)
deseq2.res <- results(deseq2.dds)
deseq2.res <- deseq2.res[order(rownames(deseq2.res)), ]
```

```{r}
head(deseq2.res)
```

```{r}
write.table(deseq2.res, "../analyses/deseq2.res-infection.tab", sep = "\t", row.names = T, quote = FALSE, col.names = TRUE)
```

```{r}
# Count number of hits with adjusted p-value less then 0.05
dim(deseq2.res[!is.na(deseq2.res$padj) & deseq2.res$padj <= 0.05, ])
```


```{r}
tmp <- deseq2.res
# The main plot
plot(tmp$baseMean, tmp$log2FoldChange, pch=20, cex=0.45, ylim=c(-15, 15), log="x", col="darkgray",
     #main="Infection Status  (pval </= 0.05)",
     xlab="mean of normalized counts",
     ylab="Log2 Fold Change")
# Getting the significant points and plotting them again so they're a different color
tmp.sig <- deseq2.res[!is.na(deseq2.res$padj) & deseq2.res$padj <= 0.05, ]
points(tmp.sig$baseMean, tmp.sig$log2FoldChange, pch=20, cex=0.45, col="red")
# 2 FC lines
abline(h=c(-1,1), col="blue")
```

Write table 
```{r}
write.table(tmp.sig, "../analyses/DEG-infection.tab", row.names = T, quote = FALSE, sep = '\t')
```


```{bash}
head ../analyses/DEG-infection.tab
```

```{bash}
wc -l ../analyses/DEG-infection.tab
```

# Heatmap - variance stabilized data transformation - top 20 
```{r}
library(pheatmap)
```

```{r}
#transform data 
vsd <- varianceStabilizingTransformation(deseq2.dds, blind = FALSE)#variance stabilizing transoformation
rld <- rlog(deseq2.dds, blind = FALSE) #controls for rows with few counts
ntd <- normTransform(deseq2.dds) #transforms data for plotting
top20 <- order(rowMeans(counts(deseq2.dds, normalized = TRUE)),
               decreasing = TRUE)[1:20] #average counts for each gene and select top 20 most abundant genes
pheatmap(assay(ntd)[top20,])
```

Try from DESeq2 manual: ntd heatmap
```{r}
select <- order(rowMeans(counts(deseq2.dds, normalized = TRUE)),
                decreasing = TRUE)[1:20]
df <- as.data.frame(colData(deseq2.dds)[,c("condition", "type")])
pheatmap(assay(ntd)[select,], cluster_rows = FALSE, show_rownames = FALSE, cluter_cols= FALSE, annotation_col = df)
```

vsd heatmap:
```{r}
pheatmap(assay(vsd)[select,], cluster_rows = FALSE, show_rownames = FALSE, cluster_cols = FALSE, annotation_col = df)
```

rld heatmap:
```{r}
pheatmap(assay(rld)[select,], cluster_rows = FALSE, show_rownames = FALSE, cluster_cols = FALSE, annotation_col = df)
```

Heatmap of sample to sample distances:
```{r}
sampleDists <- dist(t(assay(vsd)))
```

```{r}
library("RColorBrewer")
sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- paste(vsd$condition, vsd$type, sep="-")
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors)
```

PCA 
```{r}
plotPCA(vsd, intgroup=c("condition","type"))
```

```{r}
pcaData <- plotPCA(vsd, intgroup=c("condition", "type"), returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(PC1, PC2, color=condition, shape=type)) +
  geom_point(size=3) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed()
``` 




Try to save the variance stabilized transformed data for use in other heatmaps: 
```{r}
infection_vst_degs <- getVarianceStabilizedData(deseq2.dds)
head(infection_vst_degs)
```

