---
title: "get genes from enriched processes"
output: html_document
---
Rmd to get list of genes that contribute to the 15 enriched terms. 

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
```

## Upload files:

List of DEGs (123) of infection status influenced by temperature:
```{r}
contrtemp_degs <- read.delim("../analyses/DEGlist-contrast-temp-libs8-11_counts.tab", sep = '\t')
head(contrtemp_degs)
```

Upload GO database: 
```{r}
uniprot_SPgo <- read.delim("http://owl.fish.washington.edu/halfshell/bu-alanine-wd/17-07-20/uniprot-SP-GO.sorted", sep = '\t', header = FALSE)
head(uniprot_SPgo)
```


Rename some columns: 
```{r}
colnames(uniprot_SPgo) <- c("uniprot_acc_ID", "entry_name", "review_status", "protein_name", "gene_name", "organism", "length", "V8", "V9", "V10", "V11", "gene_ontology_IDs")
head(uniprot_SPgo)
```

upload `blast` output from transcriptome v. 3.1
```{r}
blastout <- read.delim("../analyses/_blast-sep-v3.1.tab", sep = '\t', header = FALSE)
head(blastout)
```

Rename some columns so that this file can be `join`-ed with the uniprot-SP-Go:
```{r}
colnames(blastout) <- c("Trinity_ID", "V2", "uniprot_acc_ID", "entry_name", "V5", "V6", "V7", "V8", "V9", "V10", "V11", "V12", "V13", "V14")
head(blastout)
```

## Annotate the DEG list:
`join` the deg list with the blast output by "Trinity_ID"
```{r}
degs_annot <- left_join(contrtemp_degs, blastout, by = "Trinity_ID")
head(degs_annot)
```

### Get GO information for DEG list:
`join` the GO database with the annotated DEG list by "Entry"
```{r}
degs_annot_GO <- left_join(degs_annot, uniprot_SPgo, by = "entry_name")
head(degs_annot_GO)
```

Pull out list of genes to do with some of the enriched terms (https://github.com/RobertsLab/paper-crab/blob/master/analyses/contrast_temp-DAVID.txt)

lipid storage (GO:GO:0019915)
```{r}
lipid <- degs_annot_GO %>%
  filter(str_detect(gene_ontology_IDs, "GO:0019915"))
head(lipid)
```
higher in cold - It's E3 ubiquitin protein ligase!! 

head involution (GO:0008258)
```{r}
invol <- degs_annot_GO %>%
  filter(str_detect(gene_ontology_IDs, "GO:0008258"))
head(invol)
```
higher in warm --> apoptotic process --> Protein scylla (Protein regulated in development and DNA damage response 2)

antennal morphogenesis (GO:0048800)
```{r}
antenn <- degs_annot_GO %>%
  filter(str_detect(gene_ontology_IDs, "GO:0048800"))
head(antenn)
```
higher in cold: protein furry (fruit fly)

wing disc dorsal/ventral pattern (GO:0048190)
```{r}
wingdv <- degs_annot_GO %>%
  filter(str_detect(gene_ontology_IDs, "GO:0048190"))
head(wingdv)
```
E3 ubiquitin (top) higher in cold
trithorax group protein osa (bottom) higher in warm 
both drosophila

imaginal disc-derived wing morphogenesis (GO:0007476)
```{r}
imgwing <- degs_annot_GO %>%
  filter(str_detect(gene_ontology_IDs, "GO:0007476"))
head(imgwing)
```
top higher in warm, bottom higher in cold - drosophila

female germline ring canal formation (GO:0007301)
```{r}
fem <- degs_annot_GO %>%
  filter(str_detect(gene_ontology_IDs, "GO:0007301"))
head(fem)
```
higher in warm - fruit fly 
response to ethanol 

transcription DNA-templated (GO:0006351)
```{r}
trnsdna <- degs_annot_GO %>%
  filter(str_detect(gene_ontology_IDs, "GO:0006351"))
head(trnsdna)
```
3 in fruit fly

determination of adult lifespan (GO:0008340)
```{r}
adult <- degs_annot_GO %>%
  filter(str_detect(gene_ontology_IDs, "GO:0008340"))
head(adult)
```
higher in warm - fruit fly

GO:0035317~imaginal disc-derived wing hair organization
```{r}
winghair <- degs_annot_GO %>%
  filter(str_detect(gene_ontology_IDs, "GO:0035317"))
head(winghair)
```
higher in cold fruit fly 

GO:0007379~segment specification
```{r}
seg <- degs_annot_GO %>%
  filter(str_detect(gene_ontology_IDs, "GO:0007379"))
head(seg)
```
both hgiher in warm --> both fruit fly 

GO:0060439~trachea morphogenesis
```{r}
trachea <- degs_annot_GO %>%
  filter(str_detect(gene_ontology_IDs, "GO:0060439"))
head(trachea)
```

GO:0030036~actin cytoskeleton organization
```{r}
cyto <- degs_annot_GO %>%
  filter(str_detect(gene_ontology_IDs, "GO:0030036"))
head(cyto)
```
hgiher in warm --> mouse 



GO:0008586~imaginal disc-derived wing vein morphogenesis
```{r}
wingvein <- degs_annot_GO %>%
  filter(str_detect(gene_ontology_IDs, "GO:0008586"))
head(wingvein)
```
top cold - fruit fly - ubiquitin protein ligase 

