---
title: "heatmaps"
output: html_document
---
Rmd to make heatmaps for the crab project. Used normalized counts. 

### Set up R Markdown document
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install/load packages:
(uncomment "install.packages("pheatmap")" if you need to install it)
```{r}
#install.packages("pheatmap")
library(pheatmap)
```

```{r}
library(dplyr)
```

```{r}
#install.packages("RColorBrewer")
library(RColorBrewer)
```

# NEW HEATMAP ATTEMPTS!

Read in normalized counts: 
```{r}
norm_counts <- read.delim("../data/salmon.isoform.TMM.EXPR.matrix")
head(norm_counts)
```
40,035 rows of normalized count data for 24 samples


Rename columns: 
```{r}
colnames(norm_counts) <- c("Trinity_ID", "CrabH_2", "CrabE_2", "CrabI_2", "CrabC_17", "CrabB_17", "CrabG_0", "CrabF_2", "CrabE_0", "CrabG_2", "CrabC_0", "CrabI_0", "CrabD_0", "CrabD_17", "CrabB_0", "CrabC_2", "CrabB_2", "CrabF_17", "CrabA_0", "CrabH_0", "CrabA_17", "CrabD_2", "CrabE_17", "CrabA_2", "CrabF_0")
head(norm_counts)
```

# TEMPERATURE DEGs
## Heatmap of temperature DEG list across all samples
Read in Temperature DEG list (423)
```{r}
temp_degs <- read.delim("../analyses/DEG-temperature.txt", header = TRUE)
head(temp_degs)
```

`join` with norm_counts by "Trinity_ID"
```{r}
temp_degs_counts <- left_join(norm_counts, temp_degs, by = "Trinity_ID")
head(temp_degs_counts)
```

Just want the ones that match! (list of 423)
```{r}
tempdegs_counts_match <- filter(temp_degs_counts, baseMean != "NA")
head(tempdegs_counts_match)
```
Pull out just the sample count data:
```{r}
tempdegs <- select(tempdegs_counts_match, Trinity_ID, CrabA_0, CrabA_2, CrabA_17, CrabB_0, CrabB_2, CrabB_17, CrabC_0, CrabC_2, CrabC_17, CrabD_0, CrabD_2, CrabD_17, CrabE_0, CrabE_2, CrabE_17, CrabF_0, CrabF_2, CrabF_17, CrabG_0, CrabG_2, CrabH_0, CrabH_2, CrabI_0, CrabI_2)
head(tempdegs)
```

Set Trinity_ID as row names:
```{r}
rownames(tempdegs) <- tempdegs$Trinity_ID #set Trinity IDs as rownames
tempdegs.heatmap <- tempdegs[,-1] #remove redundant column
head(tempdegs.heatmap)
```

Create heatmap colors:
```{r}
heatmapReds <- RColorBrewer::brewer.pal(9, "Reds")
```

Going to remove all rows that have "0" for all samples: 
```{r}
noz_tempdegs.heatmap <- tempdegs.heatmap[rowSums(tempdegs.heatmap[, -1] >0) !=0, ]
head(noz_tempdegs.heatmap)
```
402 rows
 
```{r}
pheatmap(noz_tempdegs.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapReds, fontsize_col = 20, fontsize_row = 12)
```
Try with new color scheme so that value of 0 is white:
```{r}
heatmapRedBlue <- RColorBrewer::brewer.pal(11, "RdBu")
```

```{r}
pheatmap(noz_tempdegs.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

```{r}
pheatmap(noz_tempdegs.heatmap, scale = "row", cluster_row = FALSE, cluster_cols = TRUE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

# INFECTION DEGs
## Create heatmap with infection DEG list
Just using crabs D (uninfected) and E (infected) across time

Read in infection DEG list (772)
```{r}
infection_degs <- read.delim("../analyses/DEG-infection.txt", header = TRUE)
head(infection_degs)
```

`join` with norm_counts by "Trinity_ID"
```{r}
infection_degs_counts <- left_join(norm_counts, infection_degs, by = "Trinity_ID")
head(infection_degs_counts)
```

Just want the ones that match! (list of 772)
```{r}
infectdegs_counts_match <- filter(infection_degs_counts, baseMean != "NA")
head(infectdegs_counts_match)
```

Pull out just the sample count data:
```{r}
infectiondegs <- select(infectdegs_counts_match, Trinity_ID, CrabD_0, CrabD_2, CrabD_17, CrabE_0, CrabE_2, CrabE_17)
head(infectiondegs)
```

Set Trinity_ID as row names:
```{r}
rownames(infectiondegs) <- infectiondegs$Trinity_ID #set Trinity IDs as rownames
infectiondegs.heatmap <- infectiondegs[,-1] #remove redundant column
head(infectiondegs.heatmap)
```

I think I should try to remove rows that have 0 all the way across:
```{r}
noz_infectiondegs.heatmap <- infectiondegs.heatmap[rowSums(infectiondegs.heatmap[, -1] >0) !=0, ]
head(noz_infectiondegs.heatmap)
```
now 144 rows! 

Woo! it worked. 
```{r}
pheatmap(noz_infectiondegs.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapReds, fontsize_col = 20, fontsize_row = 12)
```
Try with heatmap Red-blue (0 value = white)
```{r}
pheatmap(noz_infectiondegs.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

```{r}
pheatmap(noz_infectiondegs.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = TRUE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

```{r}
pheatmap(noz_infectiondegs.heatmap, scale = "row", cluster_row = FALSE, cluster_cols = TRUE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```
Crab E - infected 
Crab D - uninfected

### same thing but with more crabs comparing infection over time (will show effects of temperature on infection)
Pull out just the sample count data:
```{r}
infectiondegs_allcrabs <- select(infectdegs_counts_match, Trinity_ID, CrabA_0, CrabA_2, CrabA_17, CrabB_0, CrabB_2, CrabB_17, CrabC_0, CrabC_2, CrabC_17, CrabD_0, CrabD_2, CrabD_17, CrabE_0, CrabE_2, CrabE_17, CrabF_0, CrabF_2, CrabF_17, CrabG_0, CrabG_2, CrabH_0, CrabH_2, CrabI_0, CrabI_2)
head(infectiondegs_allcrabs)
```
Set Trinity_ID as row names:
```{r}
rownames(infectiondegs_allcrabs) <- infectiondegs_allcrabs$Trinity_ID #set Trinity IDs as rownames
infectiondegs_allcrabs.heatmap <- infectiondegs_allcrabs[,-1] #remove redundant column
head(infectiondegs_allcrabs.heatmap)
```

I think I should try to remove rows that have 0 all the way across:
```{r}
noz_infectiondegs_allcrabs.heatmap <- infectiondegs_allcrabs.heatmap[rowSums(infectiondegs_allcrabs.heatmap[, -1] >0) !=0, ]
head(noz_infectiondegs_allcrabs.heatmap)
```
447 rows:

```{r}
pheatmap(noz_infectiondegs_allcrabs.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

### Try again with fewer samples
Pull out just the sample count data:
```{r}
infectiondegs_alltwodays <- select(infectdegs_counts_match, Trinity_ID, CrabA_0, CrabA_2, CrabB_0, CrabB_2, CrabC_0, CrabC_2, CrabD_0, CrabD_2, CrabE_0, CrabE_2, CrabF_0, CrabF_2, CrabG_0, CrabG_2, CrabH_0, CrabH_2, CrabI_0, CrabI_2)
head(infectiondegs_alltwodays)
```

Set Trinity_ID as row names:
```{r}
rownames(infectiondegs_alltwodays) <- infectiondegs_alltwodays$Trinity_ID #set Trinity IDs as rownames
infectiondegs_alltwodays.heatmap <-infectiondegs_alltwodays[,-1] #remove redundant column
head(infectiondegs_alltwodays.heatmap)
```

I think I should try to remove rows that have 0 all the way across:
```{r}
noz_infectiondegs_alltwodays.heatmap <- infectiondegs_alltwodays.heatmap[rowSums(infectiondegs_alltwodays.heatmap[, -1] >0) !=0, ]
head(noz_infectiondegs_alltwodays.heatmap)
```
349 rows

```{r}
pheatmap(noz_infectiondegs_alltwodays.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

### Now just ABC and GHI and E (infected cold) at two time points: 

Pull out just the sample count data:
```{r}
infectiondegs_abceghitwodays <- select(infectdegs_counts_match, Trinity_ID, CrabA_0, CrabA_2, CrabB_0, CrabB_2, CrabC_0, CrabC_2, CrabE_0, CrabE_2, CrabG_0, CrabG_2, CrabH_0, CrabH_2, CrabI_0, CrabI_2)
head(infectiondegs_abceghitwodays)
```

Set Trinity_ID as row names:
```{r}
rownames(infectiondegs_abceghitwodays) <- infectiondegs_abceghitwodays$Trinity_ID #set Trinity IDs as rownames
infectiondegs_abceghitwodays.heatmap <-infectiondegs_abceghitwodays[,-1] #remove redundant column
head(infectiondegs_abceghitwodays.heatmap)
```


I think I should try to remove rows that have 0 all the way across:
```{r}
noz_infectiondegs_abceghitwodays.heatmap <- infectiondegs_abceghitwodays.heatmap[rowSums(infectiondegs_abceghitwodays.heatmap[, -1] >0) !=0, ]
head(noz_infectiondegs_abceghitwodays.heatmap)
```
345 rows

```{r}
pheatmap(noz_infectiondegs_abceghitwodays.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

```{r}
pheatmap(noz_infectiondegs_abceghitwodays.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = TRUE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

```{r}
pheatmap(noz_infectiondegs_abceghitwodays.heatmap, scale = "row", cluster_row = FALSE, cluster_cols = TRUE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```


Try same thing but with all day 0 on left, day 2 on right: 
Pull out just the sample count data:
```{r}
infectiondegs_abceghi_twodays <- select(infectdegs_counts_match, Trinity_ID, CrabA_0, CrabB_0, CrabC_0, CrabE_0, CrabG_0, CrabH_0, CrabI_0, CrabA_2, CrabB_2, CrabC_2, CrabE_2, CrabG_2, CrabH_2, CrabI_2)
head(infectiondegs_abceghi_twodays)
```

Set Trinity_ID as row names:
```{r}
rownames(infectiondegs_abceghi_twodays) <- infectiondegs_abceghi_twodays$Trinity_ID #set Trinity IDs as rownames
infectiondegs_abceghi_twodays.heatmap <-infectiondegs_abceghi_twodays[,-1] #remove redundant column
head(infectiondegs_abceghi_twodays.heatmap)
```

I think I should try to remove rows that have 0 all the way across:
```{r}
noz_infectiondegs_abceghi_twodays.heatmap <- infectiondegs_abceghi_twodays.heatmap[rowSums(infectiondegs_abceghi_twodays.heatmap[, -1] >0) !=0, ]
head(noz_infectiondegs_abceghi_twodays.heatmap)
```
345 rows

```{r}
pheatmap(noz_infectiondegs_abceghi_twodays.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

## Now with just Day 2 all samples and infection DEGs 
Pull out just the sample count data:
```{r}
infectiondegs_alldaytwo <- select(infectdegs_counts_match, Trinity_ID, CrabA_2, CrabB_2, CrabC_2, CrabD_2, CrabE_2, CrabF_2, CrabG_2, CrabH_2, CrabI_2)
head(infectiondegs_alldaytwo)
```

Set Trinity_ID as row names:
```{r}
rownames(infectiondegs_alldaytwo) <- infectiondegs_alldaytwo$Trinity_ID #set Trinity IDs as rownames
infectiondegs_alldaytwo.heatmap <-infectiondegs_alldaytwo[,-1] #remove redundant column
head(infectiondegs_alldaytwo.heatmap)
```

I think I should try to remove rows that have 0 all the way across:
```{r}
noz_infectiondegs_alldaytwo.heatmap <- infectiondegs_alldaytwo.heatmap[rowSums(infectiondegs_alldaytwo.heatmap[, -1] >0) !=0, ]
head(noz_infectiondegs_alldaytwo.heatmap)
```
252 rows
```{r}
pheatmap(noz_infectiondegs_alldaytwo.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

## Just infected crabs at day two (exclude D and F)
Pull out just the sample count data:
```{r}
infectiondegs_infectdaytwo <- select(infectdegs_counts_match, Trinity_ID, CrabA_2, CrabB_2, CrabC_2, CrabE_2, CrabG_2, CrabH_2, CrabI_2)
head(infectiondegs_infectdaytwo)
```

Set Trinity_ID as row names:
```{r}
rownames(infectiondegs_infectdaytwo) <- infectiondegs_infectdaytwo$Trinity_ID #set Trinity IDs as rownames
infectiondegs_infectdaytwo.heatmap <-infectiondegs_infectdaytwo[,-1] #remove redundant column
head(infectiondegs_infectdaytwo.heatmap)
```

I think I should try to remove rows that have 0 all the way across:
```{r}
noz_infectiondegs_infectdaytwo.heatmap <- infectiondegs_infectdaytwo.heatmap[rowSums(infectiondegs_infectdaytwo.heatmap[, -1] >0) !=0, ]
head(noz_infectiondegs_infectdaytwo.heatmap)
```
242 rows 

```{r}
pheatmap(noz_infectiondegs_infectdaytwo.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```
Just infected crabs at day 2. ABC - ambient. E- decreased. GHI - elevated. 

cluster by columns
```{r}
pheatmap(noz_infectiondegs_infectdaytwo.heatmap, scale = "row", cluster_row = FALSE, cluster_cols = TRUE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

Cluster by both:
```{r}
pheatmap(noz_infectiondegs_infectdaytwo.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = TRUE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```


## Crab A and E all time points with infection DEGs

Pull out just the sample count data:
```{r}
infectiondegs_AvE <- select(infectdegs_counts_match, Trinity_ID, CrabA_0, CrabA_2, CrabA_17, CrabE_0, CrabE_2, CrabE_17)
head(infectiondegs_AvE)
```

Set Trinity_ID as row names:
```{r}
rownames(infectiondegs_AvE) <- infectiondegs_AvE$Trinity_ID #set Trinity IDs as rownames
infectiondegs_AvE.heatmap <-infectiondegs_AvE[,-1] #remove redundant column
head(infectiondegs_AvE.heatmap)
```

I think I should try to remove rows that have 0 all the way across:
```{r}
noz_infectiondegs_AvE.heatmap <- infectiondegs_AvE.heatmap[rowSums(infectiondegs_AvE.heatmap[, -1] >0) !=0, ]
head(noz_infectiondegs_AvE.heatmap)
```
150 rows

```{r}
pheatmap(noz_infectiondegs_AvE.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```
Crab A - ambient infected
Crab E - cold infected 

## Crab B v E all time points with infection DEGs 
Pull out just the sample count data:
```{r}
infectiondegs_BvE <- select(infectdegs_counts_match, Trinity_ID, CrabB_0, CrabB_2, CrabB_17, CrabE_0, CrabE_2, CrabE_17)
head(infectiondegs_BvE)
```

Set Trinity_ID as row names:
```{r}
rownames(infectiondegs_BvE) <- infectiondegs_BvE$Trinity_ID #set Trinity IDs as rownames
infectiondegs_BvE.heatmap <-infectiondegs_BvE[,-1] #remove redundant column
head(infectiondegs_BvE.heatmap)
```


I think I should try to remove rows that have 0 all the way across:
```{r}
noz_infectiondegs_BvE.heatmap <- infectiondegs_BvE.heatmap[rowSums(infectiondegs_BvE.heatmap[, -1] >0) !=0, ]
head(noz_infectiondegs_BvE.heatmap)
```
170 rows


```{r}
pheatmap(noz_infectiondegs_BvE.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```
Crab B ambient infected
Crab E cold infected 

## **Crab C and E all time points with infection DEGs**
_This one will be used for thesis defense_ 
Pull out just the sample count data:
```{r}
infectiondegs_CvE <- select(infectdegs_counts_match, Trinity_ID, CrabC_0, CrabC_2, CrabC_17, CrabE_0, CrabE_2, CrabE_17)
head(infectiondegs_CvE)
```

Set Trinity_ID as row names:
```{r}
rownames(infectiondegs_CvE) <- infectiondegs_CvE$Trinity_ID #set Trinity IDs as rownames
infectiondegs_CvE.heatmap <-infectiondegs_CvE[,-1] #remove redundant column
head(infectiondegs_CvE.heatmap)
```

I think I should try to remove rows that have 0 all the way across:
```{r}
noz_infectiondegs_CvE.heatmap <- infectiondegs_CvE.heatmap[rowSums(infectiondegs_CvE.heatmap[, -1] >0) !=0, ]
head(noz_infectiondegs_CvE.heatmap)
```
366 rows! 

```{r}
pheatmap(noz_infectiondegs_CvE.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

### Try to get the genes that correspond to those blue clusters

from here: https://stackoverflow.com/questions/27820158/pheatmap-in-r-how-to-get-clusters 

```{r}
#set the C v E heatmap as an object
results <-
pheatmap(noz_infectiondegs_CvE.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)


```

extract 10 clusters. 
```{r}
CvE.clust <- cbind(noz_infectiondegs_CvE.heatmap,
                   cluster = cutree(results$tree_row,
                                     k = 10))
head(CvE.clust)
```
Not super clear which cluster number applies to which visual cluster. 

Next attempt: from here - https://www.biostars.org/p/287512/ 

Use the same "results" object (the heatmap) as I did above: 
```{r}
#re-order the data used to create the heatmap in the same order as what is shown in the heatmap (top to bottom)
reorg_CvE_genes <- rownames(noz_infectiondegs_CvE.heatmap[results$tree_row[["order"]],])
head(reorg_CvE_genes)
```
366 characters. Good because C v E heatmap has 366 genes! 

```{r}
#now 'cut' the data into a pre-selected number of groups. 
#not really sure how to figure out number of groups... but there seems to be 6 big blue clusters in the Crab C and Crab E heatmap...
sixclust <- sort(cutree(results$tree_row, k=6))
head(sixclust)
```

That isn't super different from my original attempt. 

Try plotting the cluster dendrogram
```{r}
plot(results$tree_row)
abline(h=7, col = "red", lty = 2, lwd =2)
```
not useful. 

Cut the row (gene) dendrogram at a Euclidean distance dis-similarity of 8 (what does that mean?)
```{r}
euc_clust <- sort(cutree(results$tree_row, h =7))
head(euc_clust)
```
I don't think that is helpful at all. The tree is too messy. 

## Crabs GHI at day 2 with infection DEGs
Pull out just the sample count data:
```{r}
infectiondegs_ghiday2 <- select(infectdegs_counts_match, Trinity_ID, CrabG_2, CrabH_2, CrabI_2)
head(infectiondegs_ghiday2)
```

Set Trinity_ID as row names:
```{r}
rownames(infectiondegs_ghiday2) <- infectiondegs_ghiday2$Trinity_ID #set Trinity IDs as rownames
infectiondegs_ghiday2.heatmap <-infectiondegs_ghiday2[,-1] #remove redundant column
head(infectiondegs_ghiday2.heatmap)
```

I think I should try to remove rows that have 0 all the way across:
```{r}
noz_infectiondegs_ghiday2.heatmap <- infectiondegs_ghiday2.heatmap[rowSums(infectiondegs_ghiday2.heatmap[, -1] >0) !=0, ]
head(noz_infectiondegs_ghiday2.heatmap)
```
106 rows



```{r}
pheatmap(noz_infectiondegs_ghiday2.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```


### Crab C, E, and H at day 2:
Pull out just the sample count data:
```{r}
infectiondegs_cehday2 <- select(infectdegs_counts_match, Trinity_ID, CrabC_2, CrabE_2, CrabH_2)
head(infectiondegs_cehday2)
```

Set Trinity_ID as row names:
```{r}
rownames(infectiondegs_cehday2) <- infectiondegs_cehday2$Trinity_ID #set Trinity IDs as rownames
infectiondegs_cehday2.heatmap <-infectiondegs_cehday2[,-1] #remove redundant column
head(infectiondegs_cehday2.heatmap)
```

I think I should try to remove rows that have 0 all the way across:
```{r}
noz_infectiondegs_cehday2.heatmap <- infectiondegs_cehday2.heatmap[rowSums(infectiondegs_cehday2.heatmap[, -1] >0) !=0, ]
head(noz_infectiondegs_cehday2.heatmap)
```

```{r}
pheatmap(noz_infectiondegs_cehday2.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```


# ABC v GHI at day 2 DEG list
## Create heatmap using ABC v GHI DEG list (5) on all samples:

Read in ABC v GHI DEG list (5)
```{r}
ABCGHI_degs <- read.delim("../analyses/DEG-ABCGHI.txt", header = TRUE)
head(ABCGHI_degs)
```

`join` with norm_counts by "Trinity_ID"
```{r}
ABCGHI_degs_counts <- left_join(norm_counts, ABCGHI_degs, by = "Trinity_ID")
head(ABCGHI_degs_counts)
```

Just want the ones that match! (list of 5)
```{r}
ABCGHI_degs_counts_match <- filter(ABCGHI_degs_counts, baseMean != "NA")
head(ABCGHI_degs_counts_match)
```

Pull out sample columns:
```{r}
abcghidegs_allcrabs <- select(ABCGHI_degs_counts_match, Trinity_ID, CrabA_0, CrabA_2, CrabA_17, CrabB_0, CrabB_2, CrabB_17, CrabC_0, CrabC_2, CrabC_17, CrabD_0, CrabD_2, CrabD_17, CrabE_0, CrabE_2, CrabE_17, CrabF_0, CrabF_2, CrabF_17, CrabG_0, CrabG_2, CrabH_0, CrabH_2, CrabI_0, CrabI_2)
head(abcghidegs_allcrabs)
```

Set Trinity_ID as row names:
```{r}
rownames(abcghidegs_allcrabs) <- abcghidegs_allcrabs$Trinity_ID #set Trinity IDs as rownames
ABCGHIdegs_all.heatmap <- abcghidegs_allcrabs[,-1] #remove redundant column
head(ABCGHIdegs_all.heatmap)
```
TRINITY_DN1994_c0_g2_i2 --> Mitochondrial import inner membrane translocase subunit Tim23	      
https://www.uniprot.org/uniprot/O14925      

TRINITY_DN3962_c0_g1_i7 --> Probable RNA-directed DNA polymerase from transposon BS (EC 2.7.7.49) (Reverse transcriptase)	      
https://www.uniprot.org/uniprot/Q95SX7        

TIRNITY_DN1456_c0_g1_i1 --> Huntingtin-interacting protein K (Huntingtin yeast partner K)        
https://pubmed.ncbi.nlm.nih.gov/27017930/         
Negative regulator of heat shock response        
https://www.uniprot.org/uniprot/Q9NX55        

TRINITY_DN365_c0_g1_i1 --> Breakpoint cluster region protein (EC 2.7.11.1)	Bcr Kiaa3017        
https://www.uniprot.org/uniprot/Q6PAJ1       

TRINITY_DN3830_c0_g1_i1 --> NO UNIPROT-SP-GO match


```{r}
pheatmap(ABCGHIdegs_all.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapReds, fontsize_col = 20, fontsize_row = 12)
```

Try with color scheme RedBlue (0 value = white)
```{r}
pheatmap(ABCGHIdegs_all.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

```{r}
pheatmap(ABCGHIdegs_all.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = TRUE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 12, fontsize_row = 12)
```

# Same heatmap idea, but just for samples ABC and GHI across Day 0 and day 2:
```{r}
abcghidegs_twodays <- select(ABCGHI_degs_counts_match, Trinity_ID, CrabA_0, CrabA_2, CrabB_0, CrabB_2, CrabC_0, CrabC_2, CrabG_0, CrabG_2, CrabH_0, CrabH_2, CrabI_0, CrabI_2)
head(abcghidegs_twodays)
```
TRINITY_DN1994_c0_g2_i2 Mitochondrial import inner membrane translocase subunit Tim23	 
Higher in Elevated temp 

TRINITY_DN3962_c0_g1_i7 Probable RNA-directed DNA polymerase from transposon BS (EC 2.7.7.49)
Higher in elevated temp

TRINITY_DN1456_c0_g1_i1 Huntingtin-interacting protein K (Huntingtin yeast partner K)
HYPK C15orf63 HSPC136	Homo sapiens (Human)
Higher in Crab B than in Crab G 

TRINITY_DN365_c0_g1_i1 Breakpoint cluster region protein (EC 2.7.11.1)	Bcr Kiaa3017	
Higher in Ambient

TRINITY_DN3830_c0_g1_i1 NO GO info
Higher in elevated


Set Trinity_ID as row names:
```{r}
rownames(abcghidegs_twodays) <- abcghidegs_twodays$Trinity_ID #set Trinity IDs as rownames
abcghidegs_twodays.heatmap <- abcghidegs_twodays[,-1] #remove redundant column
head(abcghidegs_twodays.heatmap)
```

```{r}
pheatmap(abcghidegs_twodays.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

```{r}
pheatmap(abcghidegs_twodays.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = TRUE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapRedBlue, fontsize_col = 20, fontsize_row = 12)
```

