---
title: "indiv-crabs-timeseries"
output: html_document
---
Rmd to figure out clustering of genes over time in individual crabs. Using counts from transcriptome v 3.1 

```{r}
library(dplyr)
library(tidyverse)
library(pheatmap)
library(data.table)
library(RColorBrewer)
```

Load in data of counts from individual crab RNAseq (transcriptome v 3.1):   
```{r}
norm_counts <- read.delim("../data/salmon.isoform.TMM.EXPR.matrix")
head(norm_counts)
```

Set Trinity_ID as row names:
```{r}
rownames(norm_counts) <- norm_counts$X #set Trinity IDs as rownames
norm_counts.rown <- norm_counts[,-1] #remove redundant column
head(norm_counts.rown)
```

ROUND INTEGERS UP:
```{r}
norm_counts.rown <- round(norm_counts.rown,0)
head(norm_counts.rown)
```

Set rownames as a column again:     
```{r}
norm_cts <- tibble::rownames_to_column(norm_counts.rown, "Trinity_ID")
head(norm_cts)
```

Rename sample columns: 
```{r}
colnames(norm_cts) <- c("Trinity_ID", "CrabH_2", "CrabE_2", "CrabI_2", "CrabC_17", "CrabB_17", "CrabG_0", "CrabF_2", "CrabE_0", "CrabG_2", "CrabC_0", "CrabI_0", "CrabD_0", "CrabD_17", "CrabB_0", "CrabC_2", "CrabB_2", "CrabF_17", "CrabA_0", "CrabH_0", "CrabA_17", "CrabD_2", "CrabE_17", "CrabA_2", "CrabF_0")
head(norm_cts)
```

### Subset samples from 1 crab (crab A)
```{r}
crabA <- select(norm_cts, "Trinity_ID", "CrabA_0", "CrabA_2", "CrabA_17")
head(crabA)
```

Wayyyy too many genes for a heatmap. Filter out rows that have 0 across all three timepoints:     
```{r}
crabA_noz <- crabA[rowSums(crabA[, -1] >0) !=0, ]
head(crabA_noz)
```
Now 16,427 rows. 

Now lets remove low-count genes. Say, remove rows where counts across the three samples is less than 10. 
Sum all rows in new column:         
```{r}
crabA_noz$sums <- rowSums( crabA_noz[,2:4] )
head(crabA_noz)
```

Remove columns with sums <= 30 --> didn't work. too many rows (3,504). I troubleshooted until coming to the number 55 - 2668 rows, and heatmap worked
```{r}
crabA_highcts <- filter(crabA_noz, sums >= 55)
head(crabA_highcts)
```
Now it's 3,504 rows. 

### Make a heatmap of the list of 3504 genes:
Need to remove the "sums" column for heatmap:     
```{r}
crabA_heatmap <- select(crabA_highcts, "Trinity_ID", "CrabA_0", "CrabA_2", "CrabA_17")
head(crabA_heatmap)
```

Have to make trinity ID column into rownames:
Set Trinity_ID as row names:
```{r}
rownames(crabA_heatmap) <- crabA_heatmap$Trinity_ID #set Trinity IDs as rownames
crabA.heatmap <- crabA_heatmap[,-1] #remove redundant column
head(crabA.heatmap)
```

Transform data:
```{r}
crabA.heatmap <- sqrt(crabA.heatmap)
head(crabA.heatmap)
```

Create heatmap colors:
```{r}
heatmapPurOr <- RColorBrewer::brewer.pal(11, "PuOr")
```


```{r}
pheatmap(crabA.heatmap, scale = "row", cluster_row = TRUE, cluster_cols = FALSE, clustering_distance_rows = "euclidean", clustering_method = "average", show_rownames = FALSE, show_colnames = TRUE, treeheight_col = 80, legend = TRUE, color = heatmapPurOr, fontsize_col = 20, fontsize_row = 12)
```

```{r}
my_hclust_gene <- hclust(dist(crabA.heatmap), method = "complete")
as.dendrogram(my_hclust_gene) %>%
  plot(horiz = TRUE)
```


