---
title: "Temp DEG list with gene counts"
output: html_document
---
Rmd to `join` gene count data with temperature DEG list. 

```{r}
library(tidyverse)
library(dplyr)
```

# Read in the temperature DEG list: 


### Temperature DEG annotated list
```{r}
temp_deg_annot <- read.delim("../analyses/DEG-temperature.txt")
head(temp_deg_annot)
```

# Read in table of gene counts (from Rmd: 02-DeSeq-Temperature.Rmd: 
```{r}
counts <- read.delim("../data/kallisto-0517.isoform.counts.txt", header = T, sep = "\t")
head(counts)
```

Rename "FeatureID" --> "Trinity_ID"
```{r}
colnames(counts) <- c("Trinity_ID", "380822_cold", "380823_cold", "380824_warm", "380825_warm")
head(counts)
```

# `join` the two files based on column "Trinity_ID"
```{r}
temp_deg_counts <- left_join(temp_deg_annot, counts, by = "Trinity_ID")
head(temp_deg_counts)
```

Same number of rows (423) as the original DEG list. 

# Write out the temperature comparison DEG list with gene counts for each sample: 
```{r}
write.table(temp_deg_counts, "../analyses/temp_DEGlist-counts.tab", sep = '\t', quote = FALSE, row.names = FALSE)
```

