---
title: "get-sress-response-list"
output: html_document
---
Rmd to get a list of Trinity IDs from the transcriptome that are associated with the GOslim term 'stress response'. The list should have 
35478 rows (https://github.com/RobertsLab/paper-crab/blob/master/notebooks/transcriptome-BLAST-to-GOslim.ipynb)

```{r}
library(tidyverse)
library(dplyr)
```

### Read in the uniprot-SP-GO.sorted file from the jupyter notebook listed above:
```{r}
uniprot_sp_GO <- read.delim("http://owl.fish.washington.edu/halfshell/bu-alanine-wd/17-07-20/uniprot-SP-GO.sorted", sep = '\t', header = FALSE)
head(uniprot_sp_GO)
```
168361 rows, 12 cols

#### Rename column "V1" to "uniprot_acc_ID"
```{r}
colnames(uniprot_sp_GO) <- c("uniprot_acc_ID", "entry_name", "review_status", "protein_name", "gene_name", "organism", "length", "V8", "V9", "V10", "V11", "gene_ontology_IDs")
head(uniprot_sp_GO)
```

### Read in the `BLAST` output, but the one that is called "_blast-sep.tab" from the above jupyter notebook
```{r}
blastout <- read.delim("../analyses/BLAST-to-GOslim/_blast-sep.tab", sep = '\t', header = FALSE)
head(blastout)
```
147454 rows, 14 cols

#### Rename columnn "V3" --> "uniprot_acc_ID"
```{r}
colnames(blastout) <- c("Trinity_ID", "V2", "uniprot_acc_ID", "V4", "V5", "V6", "V7", "V8", "V9", "V10", "V11", "V12", "V13", "V14")
head(blastout)
```

### `join` blastout and uniprot_sp_GO by "uniprot_acc_ID" column:
```{r}
blastout_uniprot <- left_join(blastout, uniprot_sp_GO, by = "uniprot_acc_ID")
head(blastout_uniprot)
```
147454 rows, 25 cols

Call out number of Trinity IDs that have GO terms (stat for paper)
```{r}
GOblast <- blastout_uniprot %>%
  filter(str_detect(review_status, "reviewed"))
head(GOblast)
```

70078 of the transcriptome `blast` output hits with swissprot/uniprot have GO annotations. 

### Read in the table from the BLAST to GOslim jupter notebook linked above called "Blastquery-GOslim.tab". (file is saved locally on my computer)
```{r}
blastquery_goslim <- read.table("../analyses/BLAST-to-GOslim/Blastquery-GOslim.tab", sep = '\t')
head(blastquery_goslim)
```
1395724 rows, 4 cols

#### add column names
```{r}
colnames(blastquery_goslim) <- c("Trinity_ID", "GO_ID", "GOslim", "biological_process")
head(blastquery_goslim)
```

#### Call out the GOslim terms that have the word "stress" in them: 
```{r}
stress <- blastquery_goslim %>%
  filter(str_detect(GOslim, "stress"))
head(stress)
```
35478 rows, as expected

### `join` the stress list with the annotated blast output (blastout_uniprot)
```{r}
stress_annot <- left_join(stress, blastout_uniprot, by = "Trinity_ID")
head(stress_annot)
```
36735 rows, 28 cols. Not all stress stress response have GO terms 

Find out how many of them DO have GO annotation (stat for paper):
```{r}
GOstress <- stress_annot %>%
  filter(str_detect(review_status, "reviewed"))
head(GOstress)
```
15545 rows, 28 rows
15545 of the 36735 stress response genes have associated GO annotation

### write out transcriptome stress genes to paper-crab/analyses
```{r}
write.table(stress_annot, "../analyses/stress-response-genes.tab", sep = '\t', quote = FALSE, row.names = FALSE)
```
















